### 2012-01-27 docs/notes.tex
Implement approximate likelihood
Add section in notes on EM.

To perform Gibbs sampling for a given cluster assignment $z_i$, we need to compute the likelihood for each possible assignment of $z_i$. 
This operation can be done in parallel, and only the portion of the likelihood involving the newly assigned cluster  need to be recomputed.  Computing the likelihood is then $O(T \cdot P \cdot 4N)$ (though this computation is split into $K^2$ pieces) and must computed $N \cdot K$ times per Gibbs iteration.

@@

>>>>

To perform Gibbs sampling for a given cluster assignment $z_i$, we need to compute the likelihood for each possible assignment of $z_i$. 
This operation can be done in parallel, and only the portion of the likelihood involving the newly assigned cluster  need to be recomputed.  Computing the likelihood is then $O(T \cdot P \cdot 4N)$ (though this computation is split into $K^2$ pieces) and must computed $N \cdot K$ times per Gibbs iteration.  The most straightforward implementation has space complexity $O(N^2)$ since we must keep track of $\tau_{i,j,m}$ for each $(i,j)$ at a given $m$.

@@

\section{EM for baseline}

In our baseline model, each block $(a,b) \in [1,K] \times [1,K]$ simply has a rate $\gamma_{a,b}$.
The likelihood is
\begin{align*}
\mathcal{L}(A|\lambda) &= \prod_{m=1}^M \lambda_{i_m,j_m} \prod_{(i,j) \in \mathcal{R}} \exp\{-t_M\lambda_{ij}\} \\
& =\prod_{a=1}^K \prod_{b=1}^K \prod_{m=1}^M \gamma_{ab}^{I(z_{i_m}=a,z_{j_m}=b)} \prod_{(i,j): z_i=a,z_j=b} \exp\{-t_M\gamma_{ab}\} \\
& =\prod_{a=1}^K\prod_{b=1}^K \gamma_{ab}^{n_{ab}} \exp\{-t_Mn_an_b\gamma_{ab}\}
\end{align*}
where $n_{ab} = \sum_m I(z_{i_m}=a,z_{j_m}=b)$, $n_a = \sum_{i=1}^N I(z_i=a)$.

The loglikelihood is $$\ell(A|\lambda) = \sum_{a=1}^K \sum_{b=1}^K n_{ab} \log \gamma_{ab} - t_M n_a n_b \gamma_{ab}$$
and so $\hat{\gamma}_{ab} = \frac{n_{ab}}{t_Mn_an_b}$.
